\Large{Adversarial Variational Optimization of Non-Differentiable Simulators}

### 2017-07-18 tex/paper.tex
save

Typically in variational inference one starts with a tractable prior
$p(\bftheta)$ and a tractable likelihood $p(\bfx |\bftheta )$, thus fixing the
posterior $p(\bftheta | \bfx )$.   After optimizing $q(\bftheta | \bfpsi)$ to
approximate the posterior, one would criticize the model through the
posterior predictive distribution as advocated by Box~\citep{box1980sampling}.
Effectively, we are optimizing the variational posterior $q(\bftheta | \bfpsi)$ by
directly minimizing the  divergence of the corresponding posterior predictive
distribution $\qxpsi$ and the data distribution $p_r(\bfx)$, without evaluating
the intractable likelihood $p(\bfx | \bftheta )$ or specifying a prior.
Consequently, the proposed algorithm is a form of variational inference through
model criticism~\citep{box1980sampling}.

@@

More generally, likelihood-free
inference is intimately tied to a class of algorithms that can be framed as
density estimation-by-comparison, as reviewed in \cite{2016arXiv161003483M}. In most cases, these
inference algorithms are formulated as an iterative two-step process where the
model distribution is first compared to the true data distribution and then
updated to make it more comparable to the latter. Relevant work in this
direction includes those that rely on a classifier to estimate the discrepancy betwen the observed data and the model distributions
\citep{gutmann2012noise,cranmer2015approximating,cranmer2016experiments,2016arXiv161110242D,gutmann2017likelihood,rosca2017variational}.
Also of direct relevance in the likelihood-free setup, Hamiltonian ABC~\citep{meeds2015hamiltonian}
estimate gradients with respect
to $\bftheta$ through finite differences from multiple forward passes of the
simulator with variance reduction strategies based on controlling the source of
randomness used for the latent variable $\bfz$.

>>>>

Typically in variational inference one starts with a tractable prior
$p(\bftheta)$ and a tractable likelihood $p(\bfx |\bftheta )$, thus fixing the
posterior $p(\bftheta | \bfx )$.   After optimizing $q(\bftheta | \bfpsi)$ to
approximate the posterior, one would criticize the model through the
posterior predictive distribution as advocated by Box~\citep{box1980sampling}.
In our case, we can identify the proposal distribution with the variational posterior
$q(\bftheta | \bfpsi)$, which we effectively optimize by
directly minimizing the  divergence of the corresponding posterior predictive
distribution $\qxpsi$ and the data distribution $p_r(\bfx)$, without evaluating
the intractable likelihood $p(\bfx | \bftheta )$ or specifying a prior.
Consequently, the proposed algorithm is a form of variational inference through
model criticism~\citep{box1980sampling}.

@@

More generally, likelihood-free
inference is intimately tied to a class of algorithms that can be framed as
density estimation-by-comparison, as reviewed in \cite{2016arXiv161003483M}. In most cases, these
inference algorithms are formulated as an iterative two-step process where the
model distribution is first compared to the true data distribution and then
updated to make it more comparable to the latter. Relevant work in this
direction includes those that rely on a classifier to estimate the discrepancy between the observed data and the model distributions
\citep{gutmann2012noise,cranmer2015approximating,cranmer2016experiments,2016arXiv161110242D,gutmann2017likelihood,rosca2017variational}.
Also of direct relevance in the likelihood-free setup, Hamiltonian ABC~\citep{meeds2015hamiltonian}
estimate gradients with respect
to $\bftheta$ through finite differences from multiple forward passes of the
simulator with variance reduction strategies based on controlling the source of
randomness used for the latent variable $\bfz$.
